---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# pcreg: Principal Component Regression for tree ring data

## Motivation

## How to Cite

```{r, echo = FALSE}
utils::citation("pcreg")
```

## Installation

The most up-to-date version can be installed from the GitHub repository using the `remotes` package. It can also be accomplished with the `devtools` package but devtools got a bit bloated and the components were split into individual packages. `remotes` is more streamlined if not engaging in lots of full package development and just wanting to install packages from online sources besides CRAN.

```{r, eval = FALSE}
# install.packages("remotes")
library(remotes)
remotes::install_github("lgsmith295/pcreg", build_vignettes = TRUE)
```

## Versioning

## How to Use

The vignette is the best place to start for understanding the use of the functions in the package. The `reconstruction` vignette goes through all the steps necessary to go from ITRDB data to a climate reconstruction using tree ring data.

```{r, eval = FALSE}
library(pcreg)
browseVignettes("pcreg")
vignette("reconstruction")
```

## Licenses

See the [DESCRIPTION](DESCRIPTION) and [LICENSE](LICENSE) files.

## Reporting Issues

Please report bugs to Github Issues <https://github.com/lgsmith295/pcreg/issues>. Search existing issues before reporting to see if it's already been reported and to check on the current status. If reporting a new issue add the "bug" label. Try to describe the problem in detail and include your OS and R `sessionInfo()`. Bugs can be addressed most efficiently if you include a reproducible example.

If there is something you would like to see added to the package, follow the same instructions as above but use the "enhancement" label option. A good description of the reasoning behind the feature request can be helpful for prioritizing issues. If you have code to fix a bug or provide and enhancement, please submit a GitHub pull request after reading our contributor guidelines and code of conduct.

## Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Acknowledgements

## References

