---
title: "Climate Reconstruction Using pcreg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Climate Reconstruction Using pcreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pcreg)
```

## Introduction

This package is for the development of nested principal component regression reconstructions from tree ring data. See the following publications for examples and detailed descriptions of the method: cook 1999, etc, etc. 

### Functions

There are three main functions used for making reconstruction estimates: load_clim(), eval_clim() and pcreg().

* load_clim() takes monthly climate data and averages or sums across selected months (-12:12). One can also choose to use individual months. Outputs include the climate variable and the ar model if prewhitening is indicated.

* eval_clim selects chronologies for inclusion in the model based on correlation with the climate variable. Output includes a dataframe of correlation coefficients for each chronology and the target climate variable, a dataframe of chronologies selected for inclusion in the model, and a dataframe containing the start and end year of each selected chronology. 

* pcreg() calculates principal compenent eigenvectors and linear regression model estimates, then makes predictions for each nest and splices them together. Output includes: climate reconstruction, model statistics and validation statistics for each nest, PCA objects for each nest, and lm objects for each nest.  
### Data

## Worked Example

## Comparison with PCreg

